---
title: "Assignment1"
author: "Emmeline Close"
date: "2025-09-04"
output: html_document
---

Link to GitHub Repository: https://github.com/emmelineclose/BIOL432_Assignment1

## Setup

Load libraries

```{r}
library(dplyr)
library(ggplot2)
```

## Running Previous R Scripts

Script that generates 100 instances of five species with generated limb lengths and widths as observed by 3 observers

```{r, echo=FALSE}
source("dataGenerato.R")
```

Script that calculates the leg volume for the previously mentioned 100 iterations of five species

```{r, echo=FALSE}
source("volumeEstimato.R")
```

## Importing Data

Load revised measurements.csv

```{r}
Measurements<-read.csv("./measurements.csv")
```

## Analyzing Data

Sort the data by species using dplyr

```{r}
Measurements_arranged<- Measurements %>%
  arrange(Organism, Observer, Volume)
Measurements_arranged
```

Table that shows the average volume for each species

```{r}
Volume_average_table <- Measurements %>%
  group_by(Organism) %>%
  summarize(Avg_volume = mean(Volume, na.rm = TRUE))

Volume_average_table
```

Table that shows the number of observations for each combination of species and observer

```{r}
Observations_table <- Measurements %>%
  count(Organism, Observer)

Observations_table
```

## Visualizing Data

Single box plot to compare key features of the distributions of volumes for each species

```{r}
ggplot(Measurements, aes(x = Organism, y = Volume)) + geom_boxplot() +
  labs(
    x = "Species",
    y = "Limb Volume",
    title = "Distribution of Limb Volume by Species"
  )
```

Multi-panel plot showing frequency histograms showing the distribution of limb volume for each species

```{r}
ggplot(Measurements, aes(x = Volume)) +
  geom_histogram(bins = 30, boundary = 0) + # same bins across panels
  facet_wrap(~ Organism, ncol = 3) + # one panel per species
  labs(x = "Limb Volume", y = "Frequency", title = "Limb Volume Distributions by Species")
```
